## Parallel Cross-Validation

Spark 2.3 supports parallelism in hyperparameter tuning. In other words, instead of training each model specification serially, you can now train them in parallel. This can be enabled by setting the parallelism parameter in ml_cross_validator() or ml_train_split_validation(). Here’s an example:

```{r library_load}
library(sparklyr)
sc <- spark_connect(master = "local", version = "2.3.0")
iris_tbl <- sdf_copy_to(sc, iris)

# Define the pipeline
labels <- c("setosa", "versicolor", "virginica")
pipeline <- ml_pipeline(sc) %>%
  ft_vector_assembler(
    c("Sepal_Width", "Sepal_Length", "Petal_Width", "Petal_Length"),
    "features"
  ) %>%
  ft_string_indexer_model("Species", "label", labels = labels) %>%
  ml_logistic_regression()

# Specify hyperparameter grid
grid <- list(
  logistic = list(
    elastic_net_param = c(0.25, 0.75),
    reg_param = c(1e-3, 1e-4)
  )
)

# Create the cross validator object
cv <- ml_cross_validator(
  sc, estimator = pipeline, estimator_param_maps = grid,
  evaluator = ml_multiclass_classification_evaluator(sc),
  num_folds = 3, parallelism = 4
)

# Train the models
cv_model <- ml_fit(cv, iris_tbl)
```

Once the models are trained, you can inspect the performance results by using the newly available helper function ml_validation_metrics():
```{r metrics}
ml_validation_metrics(cv_model)
spark_disconnect(sc)
```